Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-25 draft.tex
references about convexity + heuristic strategy

An important class of scoring rule is the one composed of weights that are a convex sequence, meaning that the difference between the weight of the first position and the weight of the second position is at least as great as the difference between the weights of the second and third positions, etc.
%\begin{align} \forall r \in \{1,\ldots,m-2\} \;\; & w_r - w_{r+1} \geq w_{r+1}-w_{r+2}  \\
%\iff & w_r - 2 w_{r+1} + w_{r+2} \geq 0 \label{eq:convexity}
%\end{align}
\begin{equation} 
\label{eq:convexity}
\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\W$ to denote the set of convex weight vectors.

@@

Max regret $\MR(x;\pprofile,W)$ is the worst-case loss of $x$. It is the loss occurred by an adversarial selection of a complete profile $\profile$ extending $\pprofile$ and a selection of $\w \in W$ to maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
\begin{align}
\MR(x; \pprofile, W) & = \max_{y \in A} \PMR(x,y; \pprofile, W)\\
& = \max_{\w \in W} \max_{\profile \in C(p)} \Regret(x, \profile, \w).
\end{align}

@@

We are interested in computing $PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worst than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are one rank lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:\commentOC{TODO only one number here} 
\begin{align} 
\label{eq:compl}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{align}

@@

Let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
\begin{prop}
%Assume convex weights.
The PMR can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W) =  
\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
=\max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} =
\max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
\end{align}
where the coefficient $ (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x})$ is the number of times $y$ is at rank $r$ minus the number of times $x$ is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{prop}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& C_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables. We leave $w_{1}$ and $w_{m}$ in the linear program just for clarity of presentation.

@@

\begin{itemize}
\item {\em Current solution strategy}: consider the solution of the minimax regret game, 
 $(x,y,\profile^{a},\w^{a})$, where $x$ is the minimax regret optimal alternative, $y$ the adversarial choice, $\w^{a}$ the weights, and $\profile^{a}$ the profile completed by the adversary.

Ask a question to compare $x$ with $y$ to an agent.

% Determine the part of ``regret'' due to uncertainty in $w$
% Determine the part of ``regret' due to uncertainty in $p$

>>>>

An important class of scoring rule is the one composed of weights that are a convex sequence \cite{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as great as the difference between the weights of the second and third positions, etc.
%\begin{align} \forall r \in \{1,\ldots,m-2\} \;\; & w_r - w_{r+1} \geq w_{r+1}-w_{r+2}  \\
%\iff & w_r - 2 w_{r+1} + w_{r+2} \geq 0 \label{eq:convexity}
%\end{align}
\begin{equation} 
\label{eq:convexity}
\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\W$ to denote the set of convex weight vectors.

@@

Max regret $\MR(x;\pprofile,W)$ is the worst-case loss of $x$. It is the loss occurred by an adversarial selection of a complete profile $\profile$ extending $\pprofile$ and a selection of $\w \in W$ to maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
\begin{align}
\MR(x; \pprofile, W) & = \max_{y \in A} \PMR(x,y; \pprofile, W)\\
& = \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x, \profile, \w).
\end{align}

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worst than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are one rank lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:\commentOC{TODO only one number here} 
\begin{align} 
\label{eq:compl}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{align}

@@

Let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
\begin{prop}
%Assume convex weights.
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W) =  
\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
=\max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} =
\max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
\end{align}
where the coefficient $ (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x})$ is the number of times $y$ is at rank $r$ minus the number of times $x$ is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{prop}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& C_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables. We leave $w_{1}$ and $w_{m}$ in the linear program just for clarity of presentation.

@@

\begin{itemize}
\item {\em Current solution strategy}: consider the solution of the minimax regret game, 
 $(x^{*},\bar{y},\bar{\profile},\bar{\w})$, where $x^{*}$ is the minimax regret optimal alternative,  $\bar{y} = \arg \max_{y \in A} \PMR(x,y; \pprofile, W)$; is the adversarial choice, and $\bar{\profile}$ and $\w$ are the instantiations\footnote{ $\bar{\profile} \!=\! \arg \max_{\profile \in C(\pprofile)} s(y; \profile, \bar{\w}) \!-\! s(x; \profile, \bar{\w})$; and $\bar{\w} \!=\! \arg \max_{w \in W} s(y; \pprofile, w) \!-\! s(x; \pprofile, w)$ .} of the weights associated with maximum regret.
 % Determine the part of ``regret'' due to uncertainty in $w$
% Determine the part of ``regret' due to uncertainty in $p$
Define: 
\begin{align}
\tau_{1} & = \min_{\w \in W} s(\bar{x}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w)  \label{eq:tau1} \\
\tau_{2} & = \min_{\profile \in C(\pprofile)} s(\bar{x},\profile, \bar{\w}) -  s(x^{*}; \profile, \bar{\w}).
\end{align}
Intuitively the quantity $\MMR - \tau_{1} $ measures the contribution to regret of our uncertainty about the weights $\w$, while $\MMR - \tau_{2} $ measures the uncertainty about the profile.
Therefore, whenever $\tau_{1} < \tau_{2}$ we ask a question to the chair, and compare $\bar{\w}$ and $\w_{\tau}$. the argmax of Eq. (\ref{eq:tau1}), componentwise to find the position $r$ that is most valuable to ask about.  
Instead, if $\tau_{1} > \tau_{2}$ we ask a question to one of the agent, with the following heuristics.
